---
title: "00_DataPrep_CS_MetRS"
author: "Suneeta Godbole"
date: "01/21/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())

library(openxlsx)
library(data.table)
library(stringr)

### SPIROMICS
infolder <- "C:/NetCo/Spiromics/CORE 6.1_Longitudinal_191118"
metabfolder <- "C:/NetCo/Metabolomics/NewMetabolonDatasets_20211018/processed"
clinfolder <- "C:/NetCo/Spiromics/CORE6.4_Clinical_20210105"
spiro_v5 <- "C:/NetCo/Spiromics/V5_CORE 3.0_Clinical_20210309"
outfolder <- "C:/NetCo/DiseaseProgression_Analysis"
modulefolder <- "C:/NetCo/DiseaseProgression_Analysis/consensus (1)"
metRSfolder <- "C:/NetCo/Metabolomic_Risk_Score"

### COPDGene
copd.infolder <- "C:/NetCo/Metabolomics"
copd.clinical.folder <- "C:/NetCo/COPDGene/suni_new_clinical/suni_new_clinical"

### Output
analytic.outfolder <- "C:/NetCo/Metabolomic_Risk_Score/metRS_CS_AnalyticFiles"
```

## Goal: Create a cross-sectional LASSO/Elastic Net Model of Metabolites that explains Lung Function Outcomes
Create clinical file with cross-sectional data and change variables for application of CS-elnet to delta variables

### SPIROMICS
```{r dataInSpiro}
spiro.metab <- readRDS(file.path(metabfolder,
                              "SPIROMICS_V1_LT20missing_knnImpute_metabolites_20211021.rds"))
metab.ids <- rownames(spiro.metab)

spiro.l <- read.csv(file.path(infolder, "CORE6_1_Longitudinal_191118.csv"), header = T, 
                    stringsAsFactors = F)

var.names <- names(spiro.l)

#### -- V5
# Get V5 data for Spiro
spiro.v5 <- read.csv(file.path(spiro_v5, "V5_CORE3_CLINICAL_210308.csv"), header = T, 
                     stringsAsFactors = F)
#Reduce to used variables
spiro.v5 <- spiro.v5[, c("SUBJID", "DATE_V5", "PCT_POST_FEV1_V5", "POST_FEV1_V5",
                         "POST_FEV1FVC_DERV_V5", "COPDSCORE_V5", "SGR_TOTALSCORE_V5_A")]

## -- POST_FEV1_V5 is equal to POST_FEV1_DERV -- Confirmed
names(spiro.v5) <- c("SUBJID", "DATE_VISIT", "PCT_POST_FEV1", "POST_FEV1_DERV", 
                     "POST_FEV1FVC_DERV", "COPDSCORE", "SGR_TOTALSCORE")
spiro.v5$VISIT <- "VISIT_5"
spiro.v5$POST_FEV1_ANNUALIZED_RATE <- NA
spiro.v5$DATE_VISIT <- as.character(strptime(spiro.v5$DATE_VISIT, format = "%m/%d/%y"))
### -- Add OBS_FEV1 V5 variable

spiro.obsFEV1 <- read.csv("C:/NetCo/Spiromics/V5_INV 2.0_20201012/csv_files/V5_SDF_INV2_201012.csv", 
                          header = T, stringsAsFactors = F)
spiro.obsFEV1 <- spiro.obsFEV1[, c("SUBJID", "SDF5C")]
names(spiro.obsFEV1)[names(spiro.obsFEV1) == "SDF5C"] <- "OBS_FEV1"

spiro.v5 <- merge(spiro.v5, spiro.obsFEV1, by = "SUBJID", all.x = T)

#Filter to metab ids
spiro.l <- spiro.l[spiro.l$SUBJID %in% metab.ids, c("SUBJID", "VISIT", "DATE_VISIT", "PCT_POST_FEV1",
                                                    "OBS_FEV1", "POST_FEV1FVC_DERV", 
                                                    "POST_FEV1_ANNUALIZED_RATE", "POST_FEV1_DERV",
                                                    "COPDSCORE", "SGR_TOTALSCORE")]
spiro.v5 <- spiro.v5[spiro.v5$SUBJID %in% metab.ids, ]

spiro.l <- rbind(spiro.l, spiro.v5)

# length(unique(spiro.l$SUBJID)) ## Check number of pt

### Add in Gender and Race vars from clinical file.
spiro.c <- read.csv(file.path(clinfolder, "Core6_4_clinical_20210105.csv"), header = T, 
                              stringsAsFactors = F)

spiro.c <- spiro.c[spiro.c$SUBJID %in% metab.ids, c("SUBJID", "SITE", "STRATUM_GOLD",  "GENDER", "RACE",
                                                    "ETHNICITY", "HT_CM_V1",
                                                    "DEM02_V1", "DEM03_V1", "DEM04_V1", 
                                                    "EMPHYSEMA_V1", "AGE_DERV_V1",
                                                    "BMI_CM_V1", "COPDSCORE_V1", "CURRENT_SMOKER_V1", 
                                                    "EMPHYSEMA_DIAGNOSED_V1", "SMOKING_PACK_YEARS_V1",
                                                    "LOG950_VIDA_V1",
                                                    "PRM_PERCENT_EMPHYSEM_TOTAL_V1", "DIABETES_DERV_V1", 
                                                    "CHRONIC_BRONCHITIS_V1", "CBC_BASOPHIL_CNT_V1",
                                                    "CBC_EOSINOPHIL_CNT_V1", "CBC_LYMPHOCYTE_CNT_V1",
                                                    "CBC_MONOCYTE_CNT_V1", "CBC_NEUTROPHIL_CNT_V1",
                                                    "COPDSCORE_V1", "PCT_POST_FEV1_ARC",
                                                    "PCT_POST_FEV1_V1", "POST_FEV1FVC_DERV_ARC",
                                                    "POST_FEV1FVC_DERV_V1", "POST_FEV1_DERV_ARC",
                                                    "POST_FEV1_DERV_V1", "SGR_TOTALSCORE_V1_A",
                                                    "SIX_MINUTE_WALK_DISTANCE_V1_A")]

spiro.c$STRATUM_GOLD <- factor(spiro.c$STRATUM_GOLD, 
                                levels = c(1,2,3,4), 
                                labels = c("Non-smokers", 
                                           "Post FEV1/FVC >= 0.7 (smokers, no COPD)", 
                                           "Mild to Moderate", 
                                           "Severe to Very Severe"))
spiro.c$SITE_cat <- spiro.c$SITE
spiro.c$SITE_cat[spiro.c$SITE %in% c("IA", "NJ", "TE", "UA", "UI")] <- "Oth"

spiro.c$SITE_cat <- factor(spiro.c$SITE_cat,
                           levels = c("CU", "JH", "LA", "MU", "SF", "UT", "WF", "Oth"), 
                       labels = c("Columbia", 
                                  "Johns Hopkins", 
                                  "UCLA", 
                                  "U of Michigan", 
                                  "UCSF", 
                                  "U of Utah", 
                                  "Wake Forest University", 
                                  "Other"))



spiro.c$SITE <- factor(spiro.c$SITE, 
                       levels = c("CU", "IA", "JH", "LA", "MU", "NJ", "SF", "TE", 
                                  "UA", "UI", "UT", "WF"), 
                       labels = c("Columbia", 
                                  "U of Iowa", 
                                  "Johns Hopkins", 
                                  "UCLA", 
                                  "U of Michigan", 
                                  "National Jewish Health", 
                                  "UCSF", 
                                  "Temple U", 
                                  "U of Alabama", 
                                  "U of Illinois", 
                                  "U of Utah", 
                                  "Wake Forest University"))

spiro.c$CURRENT_SMOKER_V1 <- factor(spiro.c$CURRENT_SMOKER_V1, 
                                     levels = c(0, 1), 
                                     labels = c("Not Current Smoker", 
                                                "Current Smoker"))

spiro.c$GENDER <- factor(spiro.c$GENDER, 
                          levels = c(1,2), 
                          labels = c("Male", "Female"))

spiro.c$ETHNICITY <- factor(spiro.c$ETHNICITY, 
                             levels = c(0,1), 
                             labels = c("Not Hispanic or Latino", 
                                        "Hispanic or Latino"))
spiro.c$RACE <- factor(spiro.c$RACE, 
                        levels = c(1:7), 
                        labels = c("Caucasian/White", "Black or African American", 
                                   "Asian", "American Indian or Alaska Native", 
                                   "Native Hawaiin or Other Pacific Islander", 
                                   "Mixed", "Missing"))

spiro.c$DIABETES_DERV_V1 <- factor(spiro.c$DIABETES_DERV_V1, 
                                    levels = c(0, 1), 
                                    labels = c("No", "Yes"))


spiro.c$smokStatus <- NA
spiro.c$smokStatus[spiro.c$STRATUM_GOLD == "Non-smokers" & 
                     spiro.c$CURRENT_SMOKER_V1 == "Not Current Smoker"] <- 0
spiro.c$smokStatus[spiro.c$STRATUM_GOLD != "Non-smokers" & 
                     spiro.c$CURRENT_SMOKER_V1 == "Not Current Smoker"] <- 1
spiro.c$smokStatus[spiro.c$STRATUM_GOLD != "Non-smokers" & 
                     spiro.c$CURRENT_SMOKER_V1 == "Current Smoker"] <- 2
spiro.c$smokStatus <- factor(spiro.c$smokStatus, 
                             levels = c(0,1,2), 
                             labels = c("Never Smoked", "Former Smoker", "Current Smoker"))

## CT emphysema variable (Labaki - 07/15/2021)
spiro.ct <- read.csv("C:/NetCo/Spiromics/core6_ct/CORE 6_CT_190903/Core6_CT_190903.csv", 
                     header = T, stringsAsFactors = F)
spiro.ct <- spiro.ct[, c("SUBJID", "TLC_PERCENT_EMPHYSEMA_V1", "MODEL_V1")]

spiro.c <- merge(spiro.c, spiro.ct, by = "SUBJID", all.x = T)

## CT PD15 variable (Bowler/Labaki 9/15/2021) Data from Katherine 
spiro.ct2 <- read.csv("C:/NetCo/Spiromics/V1_CT_TLC_INV15_190305 (1).csv", 
                      header = T, stringsAsFactors = F)
spiro.ct2 <- spiro.ct2[, c("SUBJID", "BOTH_TOT_V", "BOTH_HU15")]

spiro.c <- merge(spiro.c, spiro.ct2, by = "SUBJID", all.x = T)

#Create mesa adjusted lung density
spiro.c$mesa_sex <- ifelse(spiro.c$GENDER == "Male", 2,
                           ifelse(spiro.c$GENDER == "Female", 1, NA))

spiro.c$mesa_low_bmi <- ifelse(spiro.c$BMI_CM_V1 < 20, 2, 1)
spiro.c$mesa_low_bmi[is.na(spiro.c$BMI_CM_V1)] <- NA

spiro.c$mesa_high_bmi <- ifelse(spiro.c$BMI_CM_V1 > 30, 2, 1)
spiro.c$mesa_high_bmi[is.na(spiro.c$BMI_CM_V1)] <- NA

for(i in 1:nrow(spiro.c)){
  if (spiro.c$ETHNICITY[i] == "Hispanic or Latino"){
    spiro.c$mesa_hispanic[i] = 2
    spiro.c$mesa_black[i] = 1
    spiro.c$mesa_chinese[i] = 1
  }
  
  if (spiro.c$ETHNICITY[i] == "Not Hispanic or Latino" | is.na(spiro.c$ETHNICITY[i])){
    spiro.c$mesa_hispanic[i] = 1
    spiro.c$mesa_black[i] <- ifelse(spiro.c$RACE[i] == "Black or African American", 2, 1)
    spiro.c$mesa_chinese[i] <-ifelse(spiro.c$RACE[i] == "Asian", 2, 1) 
    }
}

spiro.c$TLV_pred_MESA <- (exp(7.939))*(spiro.c$AGE_DERV_V1**-0.145)*((spiro.c$HT_CM_V1/100)**2.285)*(spiro.c$mesa_sex**0.180)*(spiro.c$mesa_black**-0.241)*(spiro.c$mesa_hispanic**-0.157)*(spiro.c$mesa_chinese**-0.233)*(spiro.c$mesa_low_bmi**0.069)*(spiro.c$mesa_high_bmi**-0.086)

spiro.c$Density <- spiro.c$BOTH_HU15+1000

for(i in 1:nrow(spiro.c)){
  if(spiro.c$Density[i] > 0 & spiro.c$Density[i] <= 1000 & !(is.na(spiro.c$Density[i]))){
    spiro.c$adj_perc15[i] <- spiro.c$Density[i]*(spiro.c$BOTH_TOT_V[i]/spiro.c$TLV_pred_MESA[i])
  }
}

rownames(spiro.c) <- spiro.c$SUBJID
spiro.c <- spiro.c[order(rownames(spiro.c)), ]

spiro.c$MODEL_V1_cat <- spiro.c$MODEL_V1
spiro.c$MODEL_V1_cat[spiro.c$MODEL_V1_cat %in% c("", "Definition", "Sensation 64", "Definition AS+")] <- "Other"
```

### SPIROMICS Metabolites
Establish names for the metabolites
```{r metab_names}
var.key <- readRDS(file.path(metabfolder, "SPIROMICS_V1_MetaboliteInformation_20211021.rds"))

spiro.metab.pa <- readRDS(file.path(metabfolder,
                                    "SPIROMICS_V1_2080missing_metabolites_20211021.rds"))

cont.metabs <- names(spiro.metab)
cat.metabs <- names(spiro.metab.pa)
```

### Metabolite matching between COPDGene and SPIROMICS
```{r matchedMetabs}
## Subset to metabolites in both COPDGene and SPIROMICS
match.metab <- readRDS(file.path(metabfolder, "MatchedMetabolites.rds"))

match.metab <- match.metab[match.metab$matched.metab == 1, ] ## Select only metabs that match in type across studies

cont.matched.metab <- match.metab$metab_id[match.metab$copdg_imputation == "KNN"]
cat.matched.metab <- match.metab$metab_id[match.metab$copdg_imputation == "Present/Absent"]

spiro.metab.matched <- spiro.metab[, cont.matched.metab]
spiro.metab.pa.matched <- spiro.metab.pa[, cat.matched.metab]

met_log <- log(spiro.metab.matched) #log-transform continuous metabolites

met_all <- as.matrix(cbind(met_log, spiro.metab.pa.matched))

cormat <-  cor(met_all, method = "pearson") - diag(rep(1, ncol(met_all)))
#table(cormat > 0.8) 
#sum(abs(cormat) > 0.8)/2 ### 342 metabolites with correlations > 0.8

met_all <- met_all[order(rownames(met_all)), ]
met_all <- as.data.frame(met_all)

spiro.metab.matched <- spiro.metab.matched[order(rownames(spiro.metab.matched)), ]
spiro.metab.pa.matched <- spiro.metab.pa.matched[order(rownames(spiro.metab.pa.matched)), ]
```

### COPDGene Data
```{r readIn}
met2_no_out <- readRDS(file.path(metabfolder, "COPDGene_P2_LT20miss_knnImp_NoOut_metabs_20211021.rds"))
metab.info <- readRDS(file.path(metabfolder, "COPDGene_P2_MetaboliteInformation_20211021.rds"))
metabo.pa <- readRDS(file.path(metabfolder, "COPDGene_P2_2080miss_NoOut_metabs_20211021.rds"))
subid.keep <- rownames(met2_no_out)
```


### COPDGene Metabolites with < 20% -- log transform
```{r miss_lt20}
met_log <- log(met2_no_out)

sum(rownames(met_log) == rownames(metabo.pa))

met_copd <- cbind(met_log, metabo.pa)
met_copd <- as.data.frame(met_copd)
met_copd$sid <- rownames(met_copd)

metRS_copd <- met_copd
metRS_copd$sid <- NULL
metRS_copd <- metRS_copd[, c(colnames(met_all))]
```

### COPDGene Clinical data
```{r clin}
copd.clin <- read.csv(file.path(copd.clinical.folder, "COPDGene_P1P2P3_SM_NS_Long_Mar20.csv"), 
                      header = T, stringsAsFactors = F)
names(copd.clin)[names(copd.clin) == "ï..sid"] <- "sid"

copd.ids.P1 <- copd.clin$sid[copd.clin$visitnum == 1]
copd.ids.P2 <- copd.clin$sid[copd.clin$visitnum == 2]
copd.ids.P3 <- copd.clin$sid[copd.clin$visitnum == 3]

copd.clin <- copd.clin[copd.clin$sid %in% subid.keep, ] ## subset to pt with metabolite data

length(unique(copd.clin$sid)) ## 1125 subjects

copd.clin.p2 <- copd.clin[copd.clin$visitnum == 2, ]

## Check for variables in data set
sum(names(copd.clin.p2) %in% c("pctEmph_Thirona", "PRM_pct_emphysema_Thirona", "perc15_density", 
                               "scannerId", "Diabetes"))
```

Descriptives for COPD Gene P2 visit
```{r describeP2}
copd.clin.p2$finalgold_visit <- factor(copd.clin.p2$finalgold_visit, 
                                levels = c(-2, -1, 0, 1,2,3,4), 
                                labels = c("Never-smoked Normal",
                                           "PRISm (FEV1/FVC >= 0.7 but FEV1 < 80%)", 
                                           "GOLD 0 Control (FEV1 >= 80%, FEV1/FVC >= 0.7)", 
                                           "GOLD 1 (FEV1 >= 80%, FEV1/FVC < 0.7)", 
                                           "GOLD 2 (50% <= FEV1 < 80%, FEV1/FVC < 0.7)", 
                                           "GOLD 3 (30% <= FEV1 < 50%, FEV1/FVC < 0.7)", 
                                           "GOLD 4 (FEV1 < 30%, FEV1/FVC < 0.7)"))

copd.clin.p2$smoking_status <- factor(copd.clin.p2$smoking_status, 
                                     levels = c(0, 1, 2), 
                                     labels = c("Never Smoked", 
                                                "Former Smoker", 
                                                "Current Smoker"))

copd.clin.p2$Income <- factor(copd.clin.p2$Income,
                            levels = c(1:6), 
                            labels = c("< $15k", "$15k-$35k", "$35k-$50k", "$50k-$75k", "$75k or more",
                                       "Declines to answer"))



copd.clin.p2$EmphDxByDr <- factor(copd.clin.p2$EmphDxByDr, 
                                          levels = c(0,1), 
                                          labels = c("No", "Yes"))

copd.clin.p2$gender <- factor(copd.clin.p2$gender, 
                          levels = c(1,2), 
                          labels = c("Male", "Female"))

copd.clin.p2$ethnic <- factor(copd.clin.p2$ethnic, 
                             levels = c(1, 2), 
                             labels = c("Hispanic or Latino", 
                                        "Not Hispanic or Latino"))

copd.clin.p2$race <- factor(copd.clin.p2$race, 
                        levels = c(1:7), 
                        labels = c("Caucasian/White", "Black or African American", 
                                   "Asian", "Pacific Islander", "American Indian or Alaska Native", 
                                   "> 1 race", 
                                   "Other"))
copd.clin.p2$Diabetes <- factor(copd.clin.p2$Diabetes, 
                                levels = c(0, 1), 
                                labels = c("No", "Yes"))

#Relabel output variables from COPDGene to include indicator of phase
names(copd.clin.p2)[names(copd.clin.p2) == "FEV1_post"] <- "FEV1_post.2"
names(copd.clin.p2)[names(copd.clin.p2) == "FEV1_FVC_post"] <- "FEV1_FVC_post.2"
names(copd.clin.p2)[names(copd.clin.p2) == "pctEmph_Thirona"] <- "pctEmph_Thirona.2"
names(copd.clin.p2)[names(copd.clin.p2) == "PRM_pct_emphysema_Thirona"] <- "PRM_pct_emphysema_Thirona.2"
names(copd.clin.p2)[names(copd.clin.p2) == "FEV1pp_post"] <- "FEV1pp_post.2"
```

## Table 1
```{r tab1, eval=F}
spiro.c$RACE.ETH <- 3
spiro.c$RACE.ETH[spiro.c$ETHNICITY == "Not Hispanic or Latino" & spiro.c$RACE == "Caucasian/White"] <- 1
spiro.c$RACE.ETH[spiro.c$RACE == "Black or African American"] <- 2
spiro.c$RACE.ETH <- factor(spiro.c$RACE.ETH, 
                           levels = c(1,2,3), 
                           labels = c("Non-Hispanic, White", "Black/Africian American", 
                                      "Other"))

copd.clin.p2$race.eth <- 3
copd.clin.p2$race.eth[copd.clin.p2$race == "Caucasian/White"] <- 1
copd.clin.p2$race.eth[copd.clin.p2$race == "Black or African American"] <- 2

copd.clin.p2$race.eth <- factor(copd.clin.p2$race.eth, 
                                levels = c(1,2,3), 
                                labels = c("Non-Hispanic, White", "Black/Africian American", 
                                      "Other"))

spiro.c$tab1_gold <- NA
spiro.c$tab1_gold[spiro.c$PCT_POST_FEV1_V1 < 80.0 & spiro.c$POST_FEV1FVC_DERV_V1 >= 0.70] <- -1
spiro.c$tab1_gold[spiro.c$PCT_POST_FEV1_V1 >= 80.0 & spiro.c$POST_FEV1FVC_DERV_V1 >= 0.70] <- 0
spiro.c$tab1_gold[spiro.c$PCT_POST_FEV1_V1 >= 80.0 & spiro.c$POST_FEV1FVC_DERV_V1 < 0.70] <- 1
spiro.c$tab1_gold[spiro.c$PCT_POST_FEV1_V1 >= 50.0 & spiro.c$PCT_POST_FEV1_V1 < 80.0 
                  & spiro.c$POST_FEV1FVC_DERV_V1 < 0.70] <- 2
spiro.c$tab1_gold[spiro.c$PCT_POST_FEV1_V1 >= 30.0 & spiro.c$PCT_POST_FEV1_V1 < 50.0 
                  & spiro.c$POST_FEV1FVC_DERV_V1 < 0.70] <- 3
spiro.c$tab1_gold[spiro.c$PCT_POST_FEV1_V1 < 30.0 
                  & spiro.c$POST_FEV1FVC_DERV_V1 < 0.70] <- 4
spiro.c$tab1_gold[is.na(spiro.c$PCT_POST_FEV1_V1) | is.na(spiro.c$POST_FEV1FVC_DERV_V1)] <- NA

spiro.c$tab1_gold <- factor(spiro.c$tab1_gold, 
                            levels = c(-1, 0,1,2,3,4), 
                            labels = c("PRISm", 
                                       "COPD 0", 
                                       "COPD 1", 
                                       "COPD 2", 
                                       "COPD 3", 
                                       "COPD 4"))

copd.clin.p2$tab1_gold <- NA
copd.clin.p2$tab1_gold[copd.clin.p2$FEV1pp_post.2 < 80.0 & copd.clin.p2$FEV1_FVC_post.2 >= 0.70] <- -1
copd.clin.p2$tab1_gold[copd.clin.p2$FEV1pp_post.2 >= 80.0 & copd.clin.p2$FEV1_FVC_post.2 >= 0.70] <- 0
copd.clin.p2$tab1_gold[copd.clin.p2$FEV1pp_post.2 >= 80.0 & copd.clin.p2$FEV1_FVC_post.2 < 0.70] <- 1
copd.clin.p2$tab1_gold[copd.clin.p2$FEV1pp_post.2 >= 50.0 & copd.clin.p2$FEV1pp_post.2 < 80.0 
                  & copd.clin.p2$FEV1_FVC_post.2 < 0.70] <- 2
copd.clin.p2$tab1_gold[copd.clin.p2$FEV1pp_post.2 >= 30.0 & copd.clin.p2$FEV1pp_post.2 < 50.0 
                  & copd.clin.p2$FEV1_FVC_post.2 < 0.70] <- 3
copd.clin.p2$tab1_gold[copd.clin.p2$FEV1pp_post.2 < 30.0 
                  & copd.clin.p2$FEV1_FVC_post.2 < 0.70] <- 4
copd.clin.p2$tab1_gold[is.na(copd.clin.p2$FEV1pp_post.2) | is.na(copd.clin.p2$FEV1_FVC_post.2)] <- NA

copd.clin.p2$tab1_gold <- factor(copd.clin.p2$tab1_gold, 
                            levels = c(-1, 0,1,2,3,4), 
                            labels = c("PRISm", 
                                       "COPD 0", 
                                       "COPD 1", 
                                       "COPD 2", 
                                       "COPD 3", 
                                       "COPD 4"))

spiro.tab1 <- spiro.c[, c("SUBJID", "AGE_DERV_V1", "BMI_CM_V1", "SMOKING_PACK_YEARS_V1",
                          "POST_FEV1_DERV_V1", "POST_FEV1FVC_DERV_V1", 
                          "TLC_PERCENT_EMPHYSEMA_V1", "PRM_PERCENT_EMPHYSEM_TOTAL_V1", 
                          "GENDER", "RACE.ETH", "tab1_gold", "smokStatus", "SITE_cat", "adj_perc15",
                          "PCT_POST_FEV1_V1")]
names(spiro.tab1) <- c("sid", "age", "bmi", "smokingPackYrs",
                          "fev1_post", "fev1fvc_post", 
                          "pctEmph", "PRM_pctEmph", 
                          "gender", "race.eth", "tab1_gold", "smokingStatus", "site", "MESA_adjPD15", 
                       "FEV1pp_post.2")
spiro.tab1$Study <- "SPIROMICS"

copdg.tab1 <- copd.clin.p2[, c("sid", "age_visit", "BMI", "ATS_PackYears", 
                               "FEV1_post.2", "FEV1_FVC_post.2", "pctEmph_Thirona.2",
                               "PRM_pct_emphysema_Thirona.2", "gender", 
                               "race.eth", "tab1_gold", "smoking_status", "ccenter", "adj_density_mesa", 
                               "FEV1pp_post.2")]
names(copdg.tab1) <- c("sid", "age", "bmi", "smokingPackYrs",
                          "fev1_post", "fev1fvc_post", 
                          "pctEmph", "PRM_pctEmph", 
                          "gender", "race.eth", "tab1_gold", "smokingStatus", "site", "MESA_adjPD15", 
                       "FEV1pp_post.2")
copdg.tab1$Study <- "COPDGene"

tab1.df <- rbind(spiro.tab1, copdg.tab1)

mean.sd <- function(x, data) {
  y <- c(sum(!(is.na(data[, x]))), round(mean(data[, x], na.rm = T), 3), 
         round(sd(data[, x], na.rm= T), 2))
  return(y)
}

summaries <- data.frame(n=double(0),mean=double(0), sd=double(0), varCat = character(), 
                        stringsAsFactors = F) 

varInt <- c("age", "bmi", "smokingPackYrs", "fev1_post", "fev1fvc_post", 
            "pctEmph", "PRM_pctEmph", "MESA_adjPD15", "FEV1pp_post.2")

cat.unique <- unique(tab1.df$Study)

for(j in cat.unique){
  for(i in 1:length(varInt)){
    summaries <- rbind(summaries, c(mean.sd(varInt[i], 
                                            data = tab1.df[tab1.df$Study ==j, ]),
                                    j))
  }
}

names(summaries) <- c("n", "mean", "sd", "Study")

summaries <- cbind(varInt, summaries)

summaries.w <- reshape(summaries, 
                       timevar = "Study", 
                       idvar = "varInt", 
                       direction = "wide")

for(i in 1:length(varInt)){
  summaries.w$p.val[i] <- summary(aov(as.formula(paste0(varInt[i], "~ Study")),
                                      data = tab1.df))[[1]][1, 5]
  
}

table.deconstruct <- function(x, y, data){
  tab <- as.data.frame(table(data[, x], data[, y]))
  tab$prop <- round(as.numeric(prop.table(table(data[, x], 
                                                data[, y]), 
                                          margin= 2)), 3)*100
  
  if(all(tab$Freq >0) == T){
    tab$chisq.p <- chisq.test(data[, x], data[, y])$p.value
  }
  else(tab$chisq.p <- fisher.test(data[, x], data[, y])$p.value)

  tab$var1.name <- x
  tab$var2.name <- y
  
  return(tab)
}

cat.table <- data.frame(Var1 = character(),
                        Var2 = character(),
                        Freq = numeric(),
                        prop = numeric(),
                        stringsAsFactors = F)

varInt2 <- c("gender", "race.eth", "tab1_gold", "smokingStatus")

for(i in 1:length(varInt2)){
  res <- table.deconstruct(varInt2[i], "Study", tab1.df)
  cat.table <- rbind(cat.table, res)
}

cat.tab.w <- reshape(cat.table, 
                     timevar = "Var2", 
                     idvar = c("Var1", "var1.name", "var2.name"), 
                     direction = "wide")

names(summaries.w) <- c("var1.name", "n.spiro", "est.spiro", "sd.spiro", 
                        "n.copdg", "est.copdg", "sd.copdg", "p.val")
summaries.w$cats <- ""
names(cat.tab.w) <- c("cats", "var1.name", "var2.names", "est.copdg", "sd.copdg", 
                      "chi.copdg", "est.spiro", "sd.spiro", "p.val")


tab1 <- rbind(summaries.w[, c("var1.name", "cats", "est.spiro", "sd.spiro", 
                              "est.copdg", "sd.copdg", "p.val")], 
              cat.tab.w[, c("var1.name", "cats", "est.spiro", "sd.spiro", 
                              "est.copdg", "sd.copdg", "p.val")])
rownames(tab1) <- NULL

tab1$p.string <- ifelse(tab1$p.val >= 0.05, " ", 
                        ifelse(tab1$p.val < 0.001, "***",
                               ifelse(tab1$p.val < 0.01, "**", "*")))

tab1$spiro <- paste0(tab1$est.spiro, " (", tab1$sd.spiro, ")")
tab1$copdg <- paste0(tab1$est.copdg, " (", tab1$sd.copdg, ")")

# write.csv(tab1[, c("var1.name", "cats", "spiro", "copdg", "p.val", "p.string")], 
#           file = file.path(metRSfolder, "metRS_Table1_20210928.csv"), 
#           row.names = F)
```

## Save clinical data set
```{r saveDatasets}
saveRDS(spiro.c, file.path(analytic.outfolder, "SPIROMICS_MetRS_CS_Clinical.rds"))
saveRDS(met_all, file.path(analytic.outfolder, "SPIROMICS_MetRS_CS_Metabolites.rds"))

saveRDS(copd.clin.p2, file.path(analytic.outfolder, "COPDGene_P2_MetRS_CS_Clinical.rds"))
saveRDS(metRS_copd, file.path(analytic.outfolder, "COPDGene_P2_MetRS_CS_Metabolites.rds"))
```